\documentclass[11pt]{NSF}
\sloppy

\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{draftcopy}
\usepackage{longtable}
\usepackage{hyperref}

% some definitions
 
% begin equation, itemize, etc.

\def\be{\begin{equation}}
\def\ee{\end{equation}}
\def\bi{\begin{itemize}}
\def\ei{\end{itemize}}
\def\ben{\begin{enumerate}}
\def\een{\end{enumerate}}
\def\i{\item{}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
            
\begin{document}

\begin{center}
PHYS 1406: ``Physics of sound and music"\\
Spring 2022
\end{center}

Welcome to ``Physics of sound and music"!

Please read the Course Syllabus and Course Calendar for the upcoming semester. 
These are posted on the course website:
\url{https://josephromano.github.io/PHYS1406/}
Relevant information will also be posted via Blackboard.

Given the recent spike in COVID-19 cases due to the omicron variant, 
the first 3 weeks of class will be held remotely, via zoom.  
But this is only a temporary change.  
After that time, classes will resume in normal face-to-face mode
in SC 010.

In this class, we will discuss principles of sound, acoustics, music,
and musical instruments.  We will also have a few performances by
guest musicians %faculty from the TTU School of Music
and student volunteers from class. 

You do not have to have a background in physics and music for this
course. A high school background in science and math should be
sufficient.

The following two books are required for the course:

{\em Physics 1406 Physics of Sound and Music Course Guide and Laboratory Manual}, 
by Prof.~Walter Borst, which includes the lab manual and 
is a comprehensive collection of notes, sample homework problems, 
exams, etc.  
This should be available at the bookstore in the Student Union Building.

{\em How Music Works: The Science and Psychology of Beautiful Sounds, 
from Beethoven to the Beatles and Beyond}, by John Powell.  
Unfortunately, I didn't request it in time to have it available at the bookstore.  
But you can get it from e.g., amazon.com.  
A new paperback version is \$16.49, while the kindle version is only \$5.99.

%Please sign the attendance sheet on the front desk in the classroom 
%before each class. Come on time for signing in and finding a seat.

The lab sections of the course start on the week of Feb 7th
(see the Course Calendar for details).
The lab assignments are part of the required 
{\em PHYS 1406 Course Guide and Laboratory Manual} for the class.

I am looking forward to seeing you!

\end{document}

